### 2018-01-20 article/prework.tex
(+) prototype of article

\documentclass{article}
\usepackage{fontspec}
\usepackage{polyglossia}
\usepackage{cite}
\usepackage{hyperref}

\title{Airspace configuration with Monte Carlo tree search}
\author{Us}
\date{\today}

\begin{document}
\maketitle

\bibliography{article}
\bibliographystyle{plain}
\end{document}