### 2018-01-23 article/prework.tex
(+) added ref and written content

\documentclass{article}
\usepackage{fontspec}
\usepackage{polyglossia}
\usepackage{cite}
\usepackage{hyperref}
\usepackage{unicode-math}
\usepackage{amsmath,amsthm,stmaryrd}

\title{Airspace configuration with Monte Carlo tree search}
\author{Us}
\date{\today}

\begin{document}
\maketitle

\section{Preliminaries}
\subsection{Base theory: the environment and notations}
Notations based on Markov Decision Problem's theory (taken
from~\cite{browne2012survey})
\begin{itemize}
  \item \(S\) a set of states,
  \item \(\mathcal{A}\) a set of actions,
  \item \(T(s, a, s')\) a transition model that determines the probability of
      reaching state \(s'\) if action \(a\) is applied to state \(s\),
  \item \(R(s)\) a reward function
\end{itemize}

\subsection{Monte carlo methods~\cite{browne2012survey}}
We denote by \(Q(s, a)\) the expected reward of an action. Let
\begin{itemize}
  \item \(N(s, a)\) be
    the number of times action \(a\) has been selected from state \(s\),
  \item \(N(s)\) the number of times the game has been played from state \(s\),
  \item \(\mathbb{I}_i(s, a)\) is 1 if action \(a\) was selected from state
    \(s\) on the \(i\)th playout from state \(s\) else 0 and
  \item \(z_i\) is the result of the \(i\)th simulation played.
\end{itemize}
In Monte Carlo methods, we have
\begin{equation}
  Q(s, a) = \frac{1}{N(s, a)}\sum_{i = 1}^{N(s)}\mathbb{I}_i(s, a)z_i
\end{equation}

\subsection{Monte Carlo tree search}
In the MCTS, the action \(a\) is the path that leads from the root the the best
node computed so far

\bibliography{article}
\bibliographystyle{plain}
\end{document}